---
title: "Reproducible_Research_Assignment_1"
author: "Adeel"
date: "18/07/2020"
output: html_document
---
## Loading and preprocessing the data
```{r, echo=TRUE,warnings=FALSE}
# Adding the package and data
library(lattice)
data <- read.csv("activity.csv")
data$date <- as.Date(data$date, "%Y-%m-%d")
head(data)
```

## What is mean total number of steps taken per day?
```{r, echo=TRUE}
StepsTotal <- aggregate(steps ~ date, data = data, sum, na.rm = TRUE)
```

### The histogram looks like 
```{r}
hist(StepsTotal$steps, main = "Total steps by day", xlab = "day")
```

```{r}
#mean 
mean(StepsTotal$steps)
#median
median(StepsTotal$steps)
```

## What is average daily activity pattern?

```{r}
time_series <- tapply(data$steps, data$interval, mean, na.rm = TRUE)
plot( row.names(time_series),time_series, type = "l", xlab = "5-min interval", ylab = "Average across all Days", main = "Average number of steps taken")
```
### Which 5 minute contain maximum nunver of steps
```{r}
max_interval <- which.max(time_series)
names(max_interval)

```
## Imputing missing values
### Number of missing values
```{r}

sum(is.na(data))

```
### New dataset with filled missing valUe
```{r}
data2 <- data
data2$steps <- ifelse(is.na(data$steps), mean(data$steps, na.rm = TRUE), data$steps)

```
### Histogram with filled missing value
```{r}
StepsTotal2 <- aggregate(steps ~ date, data = data2, sum, na.rm = TRUE)
hist(StepsTotal2$steps, main = "Total steps by day", xlab = "day")
```
By imputing the missing value with mean, the historgam changed

### Mean and median after imputation
```{r}
#mean 
mean(StepsTotal2$steps)
#median
median(StepsTotal2$steps)
```

## Differences in the value in weekday and weekend
```{r}
#create a vector of weekdays
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
#Use `%in%` and `weekdays` to create a logical vector
#convert to `factor` and specify the `levels/labels`
data2$wDay <- factor((weekdays(data2$date) %in% weekdays1),levels=c(FALSE, TRUE), labels=c('weekend', 'weekday')) 
stepsByDay <- aggregate(steps ~ interval + wDay, data = data2, mean)

```

### Panel plot to show the difference over weekday and weekend
```{r}

xyplot(steps ~ interval | wDay, stepsByDay, type = "l", layout = c(1, 2), xlab = "Interval", ylab = "Number of steps")

```

